---
title: "2026_SCGL_work"
author: "Jonathan Amegin"
date: "2026-05-01"
output: html_document
---

To do: 
- for each observation, create clones in the data frame that switch the "direction_from_transect" field for later duplicate analysis
- Split the cmoparison of group sizes up between seedlings, vegetative, and reproductive
- Develop probability scoring to show 'percent similarity' between observations based on weighted scores of covariates

##Setting Up

```{r message=FALSE, warning=FALSE, include=FALSE}
# tidyverse contains both 'dyplr' and 'ggplot2'
library(tidyverse)
library(janitor)
library(ggrepel)
library(uuid)

#this contains all observations of clusters of SCGL for all sites & Transects
scgl_obs <- read_csv("data_raw/distanceSampling_1.csv")
scgl_obs <- clean_names(scgl_obs)

#This contains metadata for each transect including length & direction
scgl_trans_length <- read_csv("data_raw/transect_2.csv")
scgl_trans_length <- clean_names(scgl_trans_length)

#filtering NA's and outliers, and chaging some variable classes for later work
scgl <- scgl_obs[!(is.na(scgl_obs$transect_meter_mark|scgl_obs$distance_to_group)),] %>%  
        filter(transect_meter_mark < 110, distance_to_group < 10) %>% 
        mutate( object_id = as.factor(object_id),
                site_name = as.factor(site_name),
                transect_number = as.factor(transect_number),
                swapped_dir = 0)

scgl[is.na(scgl[,"number_seedlings_in_group"]), "number_seedlings_in_group"] <- 0
scgl[is.na(scgl[,"number_vegetative_in_group"]), "number_vegetative_in_group"] <- 0
scgl[is.na(scgl[,"number_reproductive_in_group"]), "number_reproductive_in_group"] <- 0

#adding clone observations with the 'direction_to_group' swapped for later analysis
scgl1 <- rbind(scgl, mutate(scgl, swapped_dir = "1",
                            direction_to_group = case_when(direction_to_group == "+" ~ "-",
                                                           direction_to_group == "-" ~ "+"
                                                          )))

#making new column with new UUID for each observation due to cloned entries
scgl1$new_id = UUIDgenerate(use.time = FALSE, n = nrow(scgl1), output = "string")

#use the attributes to add modifiers to the measurements            
scgl1 <- scgl1 %>% 
   mutate(
     #EW transect meter marks need to be * -1 because they were recorded as if in the(+,+) quadrant, but the mapping we will do is based on (-,+) quadrant. 
    transect_meter_mark = case_when(
      transect_orientation == "EW" ~ transect_meter_mark * -1, 
      TRUE ~ transect_meter_mark),
    #making the 'direction to group' indicate if the 'distance to group' needs to be left/right of transect.  
    distance_to_group = case_when(
      direction_to_group == "-" ~ distance_to_group * -1, 
      TRUE ~ distance_to_group ) 
          )
           
#setting up the (X,Y) columns to then transform based on plot/shapefile analysis
scgl1 <- scgl1 %>% 
  mutate( x = case_when(transect_orientation == "NS" ~ distance_to_group,
                        transect_orientation == "EW" ~ transect_meter_mark), 
          y = case_when(transect_orientation == "EW" ~ distance_to_group,
                        transect_orientation == "NS" ~ transect_meter_mark)) 

#joining the two dataframes to add transect metadata to each observation
scgl1 <- left_join(scgl1, scgl_trans_length[, c("transect_length", "global_id")],
                   by = join_by(parent_global_id == global_id))                
                
#filtering to sites
fram <- scgl1 %>% 
  filter(site_name == "12")

picnic <- scgl1 %>% 
  filter(site_name == "1"| site_name == "3")

bridgeport <- scgl1 %>% 
  filter(site_name == "10")

oilpad <- scgl1 %>% 
  filter(site_name == "11")

```

### 2026 function
Idea is to use covariates to generate a probablity score indicating likelihood of pairs of observations being duplicates. 
Covariates affecting probability:
- Number in groups, split between categories.
- Distance between observations
- Direction of transect
- within/out the crossed area

function to apply comparison to each object_id pair by referencing the fram_1 table
Needs to use the combn() pairs and parse the fram_1 table and then compare the covariates to evaluate a probability score
```{r}

test.fun <- function(site){
  #Creating list of all combinations of global_id's to reference
  list <- utils::combn(site$new_id, m = 2, simplify = FALSE)
  #Appending calculated variables to the list  
  list <- lapply(list, function(list.1){

    #indexing data with the paired global_id's
    a <- site[site$new_id == list.1[1],]
    b <- site[site$new_id == list.1[2],]

    #Calculating distance
    list.1[3] <- dist(rbind(c(a$x,a$y), c(b$x,b$y)), method = "euclidean")

    #determining if they are on the same transect
    list.1[4] <- case_when(a$transect_orientation == b$transect_orientation ~ "same",
                           a$transect_orientation != b$transect_orientation ~"different")

    #Group size comparison
    #break up into different counts 
    list.1[5] <- abs(diff(c(a$number_seedlings_in_group,b$number_seedlings_in_group)))
    list.1[6] <- abs(diff(c(a$number_vegetative_in_group,b$number_vegetative_in_group)))
    list.1[7] <- abs(diff(c(a$number_reproductive_in_group,b$number_reproductive_in_group)))
    
    #Comparing if the 'direction_from_transect' variable was swapped earlier
    #attempts to correct for potential recorder error
    #could change to a calculation to make weighting the probability score easier
    list.1[8] <- paste(a$swapped_dir, b$swapped_dir, sep = ", ")

    #using transect direction and binning distances to get probability score
    #The '500' left end of the breaks(seq()) is arbitrary as long as it is
    #larger than the furthest distance. Had trouble calculating it, so used an arbitrary large number
    #list.1[6] <-  1 / cut(as.numeric(list.1[3]), breaks = seq(0,500, 0.10), labels = FALSE)
    
     return(list.1)
  # if(z[4] == "same") 0 else

  })

  df <- data.frame("id_1"="",
                   "id_2"="",
                   "distance" = "",
                   "orientation" = "",
                   "diff_seedlings" = "",
                   "diff_vegetative"= "",
                   "diff_reproductive"= "",
                   "swapped_dir" = "")

  df <- bind_rows(lapply(list, function(z){
           data.frame("id_1" = z[1],
                      "id_2" = z[2],
                      "distance" = as.numeric(z[3]),
                      "orientation" = z[4],
                      "diff_seedlings" = z[5],
                      "diff_vegetative"= z[6],
                      "diff_reproductive"= z[7],
                      "swapped_dir" = z[8])

   }))

 return(df)
}
```

##Fram

###Plotting individuals

Transforming Fram coordinates to accurately reflect plot design.This is based on the origin (0,0) being in the lower right of the plot so that one or both of the axis matches the start of at least one transect. In this case, the origins of the three NS transects are at Y = 0, and 4 out of 5 EW transects' origins are at X = 0.

```{r fram transform}

fram_1 <- fram %>% 
  mutate( 
    x = case_when(transect_orientation == "NS" & transect_number == "1" ~ x - 10, 
                  transect_orientation == "NS" & transect_number == "2" ~ x - 20,
                  transect_orientation == "NS" & transect_number == "3" ~ x - 30,
                  transect_orientation == "EW" & transect_number == "5" ~ x - 10,
                  TRUE ~ x
                  ),
    y = case_when(transect_orientation == "EW" & transect_number == "1" ~ y + 10,
                  transect_orientation == "EW" & transect_number == "5" ~ y + 50,
                  transect_orientation == "EW" & transect_number == "7" ~ y + 70,
                  transect_orientation == "EW" & transect_number == "8" ~ y + 80,
                  transect_orientation == "EW" & transect_number == "9" ~ y + 90,
                  TRUE ~ y),
    )

#Fram NS lines tapes were laid backwards making 0 = 100. 
fram_1 <- fram_1 %>% 
  mutate( y = case_when(transect_orientation =="NS" ~ ((y - 100)* -1), TRUE ~ y)
    )

```

```{r fram graphing}
#transect start/stops for mapping
fram_lines <- tibble( x = c(-10, -20, -30, 0, -10, 0, 0, 0), 
                      y = c(0, 0, 0, 10, 50, 70, 80, 90), 
                      xend = c(-10, -20, -30, -33, -33, -60, -65, -39), 
                      yend = c(100, 100, 100, 10, 50, 70, 80, 90),
                      label = c("1NS", "2NS", "3NS", "1EW", "5EW", "7EW", "8EW", "9EW")
                    )


 fram_plot<-  ggplot( data = fram_1, mapping = aes(x = x, y = y)) +
  geom_point() +
  geom_segment(data = fram_lines, aes(x = x, y = y, xend = xend, yend = yend)) +
   geom_text(data = fram_lines,
            aes( label = label),
            nudge_x = 4,
            size = 2,
            )

  # should expect to see mirroring of each point, originals & mirrors are colored 
  fram_plot +
    geom_point(data = fram_1, aes( colour = swapped_dir))+
    coord_fixed(ratio = 1, xlim = c(-75, 10), ylim = c(0, 100))

```
```{r}
sort_by(test.fun(fram_1[sample(nrow(fram_1), size = 10),]), ~ distance)
  
```

