---
title: "2026_SCGL_work"
author: "Jonathan Amegin"
date: "2026-05-01"
output: html_document
editor_options: 
  markdown: 
    wrap: 90
---

To do:

~~- for each observation, create clones in the data frame that switch the
"direction_from_transect" field for later duplicate analysis~~

~~- Split the comparison of group sizes up between seedlings, vegetative, and
reproductive~~

~~-   Develop probability scoring to show 'percent similarity' between observations based on
    weighted scores of covariates

~~-   Keep only the duplicates with the shortest distance between the compared observations
    so that only 1 duplicate/global_id

##Setting Up

```{r message=FALSE, warning=FALSE, include=FALSE}
# tidyverse contains both 'dyplr' and 'ggplot2'
library(tidyverse)
library(janitor)
library(uuid)

#this contains all observations of clusters of SCGL for all sites & Transects
scgl_obs <- read_csv("data_raw/distanceSampling_1.csv")
scgl_obs <- clean_names(scgl_obs)

#This contains metadata for each transect including length & direction
scgl_trans_length <- read_csv("data_raw/transect_2.csv")
scgl_trans_length <- clean_names(scgl_trans_length)

#filtering NA's and outliers, and changing some variable classes for later work
scgl <- scgl_obs[!(is.na(scgl_obs$transect_meter_mark|scgl_obs$distance_to_group)),] %>%  
        filter(transect_meter_mark < 110, distance_to_group < 10) %>% 
        mutate( object_id = as.factor(object_id),
                site_name = as.factor(site_name),
                transect_number = as.factor(transect_number),
                swapped_dir = 0)

scgl[is.na(scgl[,"number_seedlings_in_group"]), "number_seedlings_in_group"] <- 0
scgl[is.na(scgl[,"number_vegetative_in_group"]), "number_vegetative_in_group"] <- 0
scgl[is.na(scgl[,"number_reproductive_in_group"]), "number_reproductive_in_group"] <- 0

#adding clone observations with the 'direction_to_group' swapped for later analysis
#some observations had NA, so need to check both directons for them
scgl1 <- rbind(scgl,
               mutate(scgl,
                      global_id = paste(global_id,"_mirror", sep = ""),
                      swapped_dir = "1",
                      direction_to_group = case_when(direction_to_group == "+" ~ "-",
                                                     direction_to_group == "-" ~ "+",
                                                     is.na(direction_to_group) ~ "NA-"
                                                     )))

scgl1[is.na(scgl1[,"direction_to_group"]), "direction_to_group"] <- "NA+"

#use the attributes to add modifiers to the measurements
#EW transect meter marks need to be * -1 because they were recorded as if in the(+,+) quadrant, but the mapping we will do is based on (-,+) quadrant.             
scgl1 <-
  scgl1 %>%
  mutate(
    transect_meter_mark = case_when(
      transect_orientation == "EW" ~ transect_meter_mark * -1,
                              TRUE ~ transect_meter_mark),
    #making the 'direction to group' indicate if the 'distance to group' needs to be left/right of transect.  
    distance_to_group = case_when(
      direction_to_group == "-" | direction_to_group == "NA-" ~ distance_to_group * -1,
                                                         TRUE ~ distance_to_group
      ))
           
#setting up the (X,Y) columns to then transform based on plot/shapefile analysis
scgl1 <- scgl1 %>% 
  mutate( x = case_when(transect_orientation == "NS" ~ distance_to_group,
                        transect_orientation == "EW" ~ transect_meter_mark), 
          y = case_when(transect_orientation == "EW" ~ distance_to_group,
                        transect_orientation == "NS" ~ transect_meter_mark)) 

#joining the two dataframes to add transect metadata to each observation
scgl1 <- left_join(scgl1, scgl_trans_length[, c("transect_length", "global_id")],
                   by = join_by(parent_global_id == global_id))                
                
#filtering to sites
fram <- scgl1 %>% 
  filter(site_name == "12")

picnic <- scgl1 %>% 
  filter(site_name == "1"| site_name == "3")

bridgeport <- scgl1 %>% 
  filter(site_name == "10")

oilpad <- scgl1 %>% 
  filter(site_name == "11")

```

### 2026 function

Idea is to use covariates to generate a probablity score indicating likelihood of pairs of
observations being duplicates.

Covariates affecting probability:

-   Number in groups, split between categories.

-   Distance between observations

-   Direction of transect

-   within/out the crossed area

function to apply comparison to each object_id pair by referencing the fram_1 table Needs
to use the combn() pairs and parse the fram_1 table and then compare the covariates to
evaluate a probability score

```{r}
###Custom Function 1
dup_fun <- function(site, distance, cluster){
  
  #making new column with new UUID for each observation due to cloned entries
  #site$new_id = UUIDgenerate(use.time = FALSE, n = nrow(site), output = "string")
  #Creating list of all combinations of global_id's to reference
  list <- utils::combn(site$global_id, m = 2, simplify = FALSE)
  #Appending calculated variables to the list  
  list <- lapply(list, function(list_1){

    #indexing data with the paired global_id's
    a <- site[site$global_id == list_1[1],]
    b <- site[site$global_id == list_1[2],]

    #Calculating distance
    list_1[3] <- dist(rbind(c(a$x,a$y), c(b$x,b$y)), method = "euclidean")

    #determining if they are on the same transect
    list_1[4] <- case_when(a$transect_orientation == b$transect_orientation ~ 1,
                           a$transect_orientation != b$transect_orientation ~ 0)

    #Group size comparisons
    list_1[5] <- abs(diff(c(a$number_seedlings_in_group,b$number_seedlings_in_group)))
    list_1[6] <- abs(diff(c(a$number_vegetative_in_group,b$number_vegetative_in_group)))
    list_1[7] <- abs(diff(c(a$number_reproductive_in_group,b$number_reproductive_in_group)))
    
    #Comparing if the 'direction_from_transect' variable was swapped earlier
    #attempts to correct for potential recorder error
    #could change to a calculation to make weighting the probability score easier
    list_1[8] <- as.numeric(a$swapped_dir) + as.numeric(b$swapped_dir)
    list_1[9] <- case_when(str_detect(a$direction_to_group, pattern = "NA") &
                            str_detect(b$direction_to_group, pattern = "NA") ~ 2,
                           xor(str_detect(a$direction_to_group, pattern = "NA"), 
                            str_detect(b$direction_to_group, pattern = "NA")) ~ 1,
                            TRUE ~ 0)
     return(list_1)
  })
  
  #use the hamilton, et. al discrepancy score equation as a base
  list <- lapply(list, function(z){
    z[10] <- 
      (min(as.numeric(z[3]), distance)/ distance) + # using 0.5m as the 'beyond a doubt' not duplicate distance
      (min(as.numeric(z[5]), 8)/8) + #seedlings difference, 8 = not duplicate
      (min(as.numeric(z[6]), 5)/5) + #vegetative difference, 5 = not duplicate
      (min(as.numeric(z[7]), 2)/2) #reproductive difference, 2 = not duplicate
    
    z[11] <-  #fixed tolerance test
      if((as.numeric(z[3]) < distance) & (sum(as.numeric(z[5],z[6],z[7])) < cluster)) 1 else 0
    
    z[12] <- if((as.numeric(z[3]) < distance)) 1 else 0
                
    return(z)
  })

  df <- bind_rows(lapply(list, function(z){
           data.frame("id_1" = z[1],
                      "id_2" = z[2],
                      "distance" = as.numeric(z[3]),
                      "orientation" = z[4],
                      "diff_seedlings" = z[5],
                      "diff_vegetative"= z[6],
                      "diff_reproductive"= z[7],
                      "swapped_dir" = z[8],
                      "NA_dir" = z[9],
                      "discrepency" = as.numeric(z[10]),
                      "fixed_tolerance" = z[11],
                      "fixed_distance" = z[12])
   }))
 df$dup_id <- UUIDgenerate(use.time = FALSE, n = nrow(df), output = "string")
 
 return(df)
}
```

```{r}
####Custom Function 2
#checking for observations that have multiple duplicates & then filtering down to 'most likely'
dup_fun_1 <-
  function(site){
  df <- filter(site, discrepency <1 & orientation == 0) #filter to possible dupes
  id <- c(df$id_1, df$id_2) #vector of all id's that are associated with duplicated observations
  if (sum(duplicated(id)) != 0){
    dup_id <- unique(id[duplicated(id)])
    id_1 <- #function to find the multiples that are NOT the most likely
    lapply(dup_id, function(i){
        df_1 <- filter(df, id_1 == i | id_2 == i)
        df_2 <- filter(df_1, discrepency != min(discrepency))$dup_id
        df_2
    })
    df[-which(df$dup_id %in% unlist(id_1)),]}
  else{df} 
}

```

```{r appending df with dup ids}
join_fun <- 
function(site, d, c){
site_1 <- filter(site, zoned == 1)
  # site.ew <- filter(site.1, transect_orientation == "EW" & swapped_dir == 0)
  # site.ns <- filter(site.1, transect_orientation == "NS" & swapped_dir == 0)
  site_1d <- dup_fun(site_1, distance = d, cluster = c)
  #getting the numbers of duplicates for both the discrepency score & fixed tolerance 
  #analysis allows for comparison between methods
  site_score <-   dup_fun_1(filter(site_1d, discrepency < 1 & swapped_dir == 0))
  site_score_1 <- dup_fun_1(filter(site_1d, discrepency < 1 & swapped_dir < 2))
  site_score_2 <- dup_fun_1(filter(site_1d, discrepency < 1 & swapped_dir <= 2))
  
  site_fixed <-   dup_fun_1(filter(site_1d, fixed_tolerance == 1 & swapped_dir == 0))
  site_fixed_1 <- dup_fun_1(filter(site_1d, fixed_tolerance == 1 & swapped_dir < 2))
  site_fixed_2 <- dup_fun_1(filter(site_1d, fixed_tolerance == 1 & swapped_dir <= 2))
  
  site_distance <-   dup_fun_1(filter(site_1d, fixed_distance == 1 & swapped_dir == 0))
  site_distance_1 <- dup_fun_1(filter(site_1d, fixed_distance == 1 & swapped_dir < 2))
  site_distance_2 <- dup_fun_1(filter(site_1d, fixed_distance == 1 & swapped_dir <= 2))
  
  left_join(site, select(site_score, id_1, dup_id), by = join_by(global_id == id_1)) %>% 
    rename(dup_id_1 = dup_id) %>% 
    left_join(select(site_score, id_2, dup_id), by = join_by(global_id == id_2)) %>% 
    unite("site_score", c("dup_id_1", "dup_id"), na.rm = TRUE) %>% 
  
  left_join( select(site_score_1, id_1, dup_id), by = join_by(global_id == id_1)) %>% 
    rename(dup_id_1 = dup_id) %>% 
    left_join(select(site_score_1, id_2, dup_id), by = join_by(global_id == id_2)) %>% 
    unite("site_score_1", c("dup_id_1", "dup_id"), na.rm = TRUE) %>% 
  
  left_join( select(site_score_2, id_1, dup_id), by = join_by(global_id == id_1)) %>% 
    rename(dup_id_1 = dup_id) %>% 
    left_join(select(site_score_2, id_2, dup_id), by = join_by(global_id == id_2)) %>% 
    unite("site_score_2", c("dup_id_1", "dup_id"), na.rm = TRUE) %>% 
  
  left_join( select(site_fixed, id_1, dup_id), by = join_by(global_id == id_1)) %>% 
    rename(dup_id_1 = dup_id) %>% 
    left_join(select(site_fixed, id_2, dup_id), by = join_by(global_id == id_2)) %>% 
    unite("site_fixed", c("dup_id_1", "dup_id"), na.rm = TRUE) %>% 
  
  left_join( select(site_fixed_1, id_1, dup_id), by = join_by(global_id == id_1)) %>% 
    rename(dup_id_1 = dup_id) %>% 
    left_join(select(site_fixed_1, id_2, dup_id), by = join_by(global_id == id_2)) %>% 
    unite("site_fixed_1", c("dup_id_1", "dup_id"), na.rm = TRUE) %>% 
  
  left_join( select(site_fixed_2, id_1, dup_id), by = join_by(global_id == id_1)) %>% 
    rename(dup_id_1 = dup_id) %>% 
    left_join(select(site_fixed_2, id_2, dup_id), by = join_by(global_id == id_2)) %>% 
    unite("site_fixed_2", c("dup_id_1", "dup_id"), na.rm = TRUE) %>% 
  
  left_join( select(site_distance, id_1, dup_id), by = join_by(global_id == id_1)) %>% 
    rename(dup_id_1 = dup_id) %>% 
    left_join(select(site_distance, id_2, dup_id), by = join_by(global_id == id_2)) %>% 
    unite("site_distance", c("dup_id_1", "dup_id"), na.rm = TRUE) %>% 
  
  left_join( select(site_distance_1, id_1, dup_id), by = join_by(global_id == id_1)) %>% 
    rename(dup_id_1 = dup_id) %>% 
    left_join(select(site_distance_1, id_2, dup_id), by = join_by(global_id == id_2)) %>% 
    unite("site_distance_1", c("dup_id_1", "dup_id"), na.rm = TRUE) %>% 
  
  left_join( select(site_distance_2, id_1, dup_id), by = join_by(global_id == id_1)) %>% 
    rename(dup_id_1 = dup_id) %>% 
    left_join(select(site_distance_2, id_2, dup_id), by = join_by(global_id == id_2)) %>% 
    unite("site_distance_2", c("dup_id_1", "dup_id"), na.rm = TRUE)
  
} 

```

##picnic
```{r}
#transforming coordinates to match the plot design. This requires accurate note taking
 picnic_1 <-  picnic %>% 
   mutate( 
    x = case_when(transect_orientation == "NS" & transect_number == "1" ~ x - 20, 
                  transect_orientation == "NS" & transect_number == "2" ~ x - 30,
                  transect_orientation == "NS" & transect_number == "3" ~ x - 40,
                  transect_orientation == "EW" & transect_number == "1" ~ x - 20,
                  transect_orientation == "EW" & transect_number == "2" ~ x - 10,
                  transect_orientation == "EW" & transect_number == "3" ~ x - 10,
                  transect_orientation == "EW" & transect_number == "4" ~ x
                  ),
    y = case_when(transect_orientation == "NS" & transect_number == "1" ~ y + 20, 
                  transect_orientation == "NS" & transect_number == "2" ~ y + 10,
                  transect_orientation == "NS" & transect_number == "3" ~ y ,
                  transect_orientation == "EW" & transect_number == "1" ~ y + 20,
                  transect_orientation == "EW" & transect_number == "2" ~ y + 30,
                  transect_orientation == "EW" & transect_number == "3" ~ y + 40,
                  transect_orientation == "EW" & transect_number == "4" ~ y + 50,
                  ))

#adding an attribute to show if an observation is within the zone that NS & EW transects share
picnic_1 <-
  mutate(picnic_1,
         zoned = case_when(
           c(rowSums(mapply(FUN = function(a, b){between(a, b-3, b+3)}, picnic_1[, "y"], c(20, 30, 40, 50))) +
             rowSums(mapply(FUN = function(a, b){between(a, b-3, b+3)}, picnic_1[, "x"], c(-20, -30, -40)))) == 2 ~ 1,
           TRUE ~ 0))

picnic_dup <- join_fun(site = picnic_1, d = 1, c = 10)

picnic_dup
 
```

```{r}
#Graphing the site with duplicated observations marked in blue

pic_lines <- tibble( x = c(-20, -30, -40, -20, -10, -10, 0), 
                     y = c(20, 10, 0, 20, 30, 40, 50), 
                     xend = c(-20, -30, -40, -90, -70, -70, -50),
                     yend = c(60, 76, 50, 20, 30, 40, 50),
                     label = c("1NS", "2NS", "3NS", "1EW", "2EW", "3EW", "4EW"))

pic_plot <- picnic_dup %>% 
  filter(zoned ==1 & swapped_dir == 0) %>% 
  ggplot( mapping = aes(x, y)) +
  geom_point(aes(color = transect_orientation), shape = 1) +
  geom_point(data = filter(picnic_dup, site_score_1 != "" & swapped_dir == 0), color = "blue") +
  geom_segment(data = pic_lines, aes(x = x, y = y, xend = xend, yend = yend))

   
pic_plot

```

