---
title: "SCGL Test R"
author: "Jonathan Amegin"
date: "2024-06-03"
output: html_document
---

# SCGL Distance Sampling data analysis

A test project cleaning and analyzing the Distance Sampling data collected on the Sclerocactus glauca (SCGL) trip

End goals:

-   Plot the individual observations for each site on a coordinate grid
-   Mark the observations which overlap/were observed both going NS & EW.
-   Get a percent observational error to apply across the observed population.

```{r message=FALSE, warning=FALSE, include=FALSE}
# tidyverse contains both 'dyplr' and 'ggplot2'
library(tidyverse)

```

There are 3 associated .csv files.

-   "distanceSampling_1.csv" contains all of the individual observations of SCGL
-   "Sclerocactus_glaucus_Distance_Sampling_0.csv" contains the distance between transects
-   "transect_2.csv" contains the transect number and transect length

*These are needed to join the transect length data into the "distanceSampling_1.csv" data*

```{r loading csv's message=FALSE, warning=FALSE, include=FALSE}
scgl_obs <- read_csv("~/SCGL_TEST_R/data/distanceSampling_1.csv")
scgl_trans_spacing <- read_csv("~/SCGL_TEST_R/data/Sclerocactus_glaucus_Distance_Sampling_0.csv")
scgl_trans_length <- read_csv("~/SCGL_TEST_R/data/transect_2.csv")
```

## Take a Look:


## Clean Up

search for any NA's in the data


