---
title: "SCGL Test R"
author: "Jonathan Amegin"
date: "2024-06-03"
output: html_document
---

# SCGL Distance Sampling data analysis

A test project cleaning and analyzing the Distance Sampling data collected on the Sclerocactus glauca (SCGL) trip

End goals:

-   Plot the individual observations for each site on a coordinate grid
-   Mark the observations which overlap/were observed both going NS & EW.
-   Get a percent observational error to apply across the observed population.

```{r message=FALSE, warning=FALSE, include=FALSE}
# tidyverse contains both 'dyplr' and 'ggplot2'
library(tidyverse)
library(dplyr)
```

There are 3 associated .csv files.

-   "distanceSampling_1.csv" contains all of the individual observations of SCGL
-   "Sclerocactus_glaucus_Distance_Sampling_0.csv" contains the distance between transects
-   "transect_2.csv" contains the transect number and transect length

*These are needed to join the transect length data into the "distanceSampling_1.csv" data*

```{r loading csv's message=FALSE, warning=FALSE, include=FALSE}
scgl_obs <- read_csv("data_raw/distanceSampling_1.csv")
scgl_trans_spacing <- read_csv("data_raw/Sclerocactus_glaucus_Distance_Sampling_0.csv")
scgl_trans_length <- read_csv("data_raw/transect_2.csv")
```

## Take a Look:


## Clean Up

'cactus' will contain our cactus observations with only the relevant data for placing the locations.
```{r}
cactus <- scgl_obs %>%  
  select(`Transect Meter Mark`, `transectNumber_`, `transectOrientation`, `siteName`, `distBetweenTransects`, `Distance To Group`)
head(cactus)

```

we can look at total number of cactus seen per site. 
```{r}
cactus %>% 
  group_by(siteName) %>% 
  summarize(total = sum(!is.na(`Transect Meter Mark`)) )
```

```{r}
cactus %>% 
  mutate( transectOrientation = factor(transectOrientation)) %>% 
  mutate( transectNumber_ = factor(transectNumber_)) %>% 
  mutate( siteName = factor(siteName))
```

```{r}
cactus %>% 
  filter(`Transect Meter Mark` > 100 & `Distance To Group` > 4)
```

```{r}
scgl_obs %>% 
filter(`Transect Meter Mark` > 100 & `Distance To Group` > 4)

```

Join the two dataframes to show the length of the transect that each cactus was observed on. Joins based on the ParentGlobalID of the cactus observation matching the GlobalID of the transect it was observed on.
Then remove several columns that will not contribute to the overall analysis.

```{r}
#make a clean version of 'scgl_trans_length' omitting extra data not useful to alaysis
clean_length <- scgl_trans_length %>% 
  select(GlobalID, `Transect Number`,`Transect Direction`,`Transect Length`)
```

```{r}
joined_scgl <- left_join(scgl_obs, clean_length, join_by ("ParentGlobalID"=="GlobalID"), relationship = "many-to-one")

#cleans up extra data we dont need for looking at just the cactus
joined_scgl <- joined_scgl[,-c(11,12,15:21)]
```

```{r}

unique(joined_scgl$siteName)

#joined_scgl$siteName <- na.omit(joined_scgl$siteName)  not quite the right idea

site_11 <- joined_scgl %>% 
  filter(siteName == "11")
# site_names <- c("site_1", "site_3","site_10", "site_11", "site_12")   want to create vector of variable names to assign new tables containing only the observations for a specific site. Think I can do it with a loop instead of writing each one out. Maybe use an anonymous function to write a looping script 
#that reads the unique variables for 'siteName' and writes an object of the filtered table based on the character string. or use name() function after filter looping based on unique variables. 
```


